{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}

### 2016-12-30 proposal/overview.tex
write overview of mutexes

% TODO

@@

\subsection{Pintos}
% TODO

>>>>

To prevent unexpected interleavings such as the example in Figure~\ref{fig:concurrency-bug}(b),
most concurrent programs use {\em concurrency primitives} to control which interleavings are possible.
Controlling nondeterminism is not typically provided by any features of programming languages themselves;
rather, it is achieved via special atomicity mechanisms provided by the CPU and/or operating system -- hence the term ``primitive''.
For example, x86 CPUs provide the {\tt xchg} instruction, which performs both a read and subsequent write to some shared memory, with no possibility for other logic to interleave in between.
Using such atomic instructions as building blocks, concurrency libraries provide abstractions for controlling nondeterminism in several commonly-desired ways.
These include {\em locks}, {\em descheduling}, {\em condition variables}, {\em semaphores}, {\em reader-writer locks}, and {\em message-passing}.

\begin{figure}[t]
	\begin{tabular}{p{0.45\textwidth}p{0.5\textwidth}}
		{\footnotesize
		\begin{tabular}{l}
			\texttt{typedef struct mutex \{} \\
			\texttt{~~~~volatile int held;} \\
			\texttt{~~~~int owner;} \\
			\texttt{\} mutex\_t;} \\
			\texttt{void mutex\_lock(mutex\_t *mp) \{}\\
			\texttt{~~~~while (xchg(mp->held, 1))} \\
			\texttt{~~~~~~~~yield(mp->owner);} \\
			\texttt{~~~~mp->owner = gettid();} \\
			\texttt{\}} \\
			\texttt{void mutex\_unlock(mutex\_t *mp) \{}\\
			\texttt{~~~~mp->owner = -1;} \\
			\texttt{~~~~mp->held = 0;} \\
			\texttt{\}} \\
		\end{tabular}
		}
		&
		{\footnotesize
		\begin{tabular}{l}
			\texttt{int x;} \\
			\texttt{mutex\_t m;} \\
			\texttt{void count() \{} \\
			\texttt{~~~~for (int i = 0; i < 1000;} \\
			\texttt{~~~~~~~~~i++)} \\
			\texttt{~~~~~~~~mutex\_lock(\&m);} \\
			\texttt{~~~~~~~~x++;} \\
			\texttt{~~~~~~~~mutex\_unlock(\&m);} \\
			\texttt{\}} \\
		\end{tabular}
		}
		\\
		(a) A simple mutual exclusion lock built using the {\tt xchg} instruction. %(accessed using a GCC compiler intrinsic).
		&
		(b) The {\tt count} function from Figure~\ref{fig:concurrency-bug}, adjusted to use a mutex to ensure each increment of {\tt x} is uninterruptible.
	\end{tabular}
	\caption{Using a locking primitive to protect accesses to shared state.}
	\label{fig:mutex}
\end{figure}

Each such abstraction provides certain semantics about what thread interleavings can arise surrounding their use.
When building a tool for testing concurrent programs,
one may include some computational understanding of the behaviour of any, or all, of these abstractions.
Annotating a certain abstraction's semantics treats it as a trusted concurrency primitive in its own right,
and allows the testing tool to reduce the possible space of interleavings (or the set of false positive data-race candidates reported, etc.),
at the cost of increasing the implementation and theoretical complexity of the analysis.
In this thesis, I will consider locks and descheduling to be the only concurrency primitives,
and assume the others listed above are implemented using those as building blocks (an exercise for the reader \cite{thrlib}).

Locks (or {\em mutexes}, short for ``mutual exclusion locks'') are objects, shared by multiple threads, which allow the programmer to mark certain {\em critical sections} of code that must not interleave with each other.
When one thread completes a call to {\tt mutex\_lock(mp)}, all invocations by other threads on the same {\tt mp} will wait (or ``block'') until the corresponding {\tt mutex\_unlock(mp)}.
Figure~\ref{fig:mutex}(a) shows how a yielding mutex (not the best implementation, but the simplest) may be implemented using {\tt xchg},
and (b) shows how a mutex may be used to fix the example from Figure~\ref{fig:concurrency-bug}.

@@

\subsection{Pintos}
\label{sec:overview-pintos}
% TODO
